project-plan

### 2025-09-26 project-plan.qmd
Comment on project plan

Among trauma-related conditions, traumatic brain injury (TBI) is a major cause of death and disability worldwide and
constitutes a significant public health burden. (3) TBI is the result of an external force causing neurological
impairments that may be either transient or permanent.  (4) Furthermore, TBI should not be regarded as a single event
but rather as a dynamic process. The initial trauma is followed by secondary pathophysiological cascades, including
ischemia, oedema, and metabolic disturbances, that develop over time and critically influence long-term outcomes. (5)
As such, patients affected often require admission to the intensive care unit (ICU).

@@

@@

(At Karolinska University Hospital, the setting for this study, similar principles are seen in the Traumahandboken.
For severe TBI, defined as GCS < 8, early airway protection with intubation in unconscious patients, rapid CT
imaging, timely administration of tranexamic acid, and early neurosurgical consultation are recommended. In cases
where there are signs of impending herniation, the handbook advises measures such as controlled hyperventilation and
osmotherapy with mannitol or hypertonic saline. All this is always in conjunction with urgent neurosurgical
evaluation. (Källa Karolinska handboken))  <!-- Remove paragraph? -->

@@

Despite the presence of systems and guidelines in place for these patients, trauma care remains vulnerable to errors,
deviations from established practice, and treatment delays, all of which may negatively influence outcomes. In
response, the American College of Surgeons developed the Trauma Quality Improvement Program (TQIP), a risk-adjusted
benchmarking system designed to assess the quality of trauma care. Through audit filters and peer review, it
identifies opportunities for improvement (OFI) and ultimately reduces trauma-related mortality (10,11). The WHO, however, insist that the trauma quality improvement programmes must be a continuous, multidisciplinary process. This
entails audit filters, structured morbidity and mortality and preventable death case reviews in order to find
opportunities for improvement (OFI). They also discuss that the formulation of corrective strategies or action plans
in response to identified OFI:s is essential to trauma quality improvement programmes. Furthermore, the WHO
emphasises that TQIP must ensure that corrective measures addressing identified OFI result in tangible and verifiable
improvements in outcomes, thereby confirming that deficiencies are not only recognised but also effectively resolved. (Källa WHO) <!-- add WHO source -->

@@

@@

Aim
=====
As such, this study aims to identify and describe opportunities for improvement in the care of adult trauma patients
with TBI, and to examine patient- and process-related factors associated with their occurrence.

@@

Study Design
-----------
A single-centre, registry-based retrospective cohort study was conducted at Karolinska University Hospital in Solna
(NKS). The trauma registry for NKS is a comprehensive trauma registry that reports to the Swedish National Trauma
Registry, SweTrau, and is compliant with the Utstein template for uniform reporting after trauma (Källa  Utstein
Template). As such, the registry prospectively records patient demographics, prehospital and emergency department
physiology, times of key interventions such as imaging, surgery and transfer, injury severity scores, and in-hospital outcome.

The Swedish Trauma Registry includes patients of all ages who have sustained serious injuries from traumatic events
such as falls, traffic incidents, or other external violence. Since 2024 , patients have been included if a trauma
team activation has occurred, regardless of injury severity, or if they have a New Injury Severity Score (NISS)
greater than fifteen without trauma team activation. Patients transferred to a hospital within seven days of the
traumatic event with a NISS greater than fifteen are also included. Exclusion criteria are cases where the only
traumatic condition is chronic subdural haematoma, trauma alarms without an underlying traumatic event, patients with
protected identity, and patients declared dead on arrival. Since these inclusion and exclusion criteria have changed
over time, patients in this study will be filtered to ensure they fulfil the criteria relevant to the study period
(Källa SweTrau årsbok 2024). <!--Use the 2023 version instead? Data used is up to 2023 -->

@@

Setting
---------
Karolinska University Hospital in Solna houses Trauma Centre Karolinska (TCK), Sweden’s largest provider of advanced
trauma care for both adults and children. According to the definition of the American College of Surgeons, Karolinska
University Hospital in Solna serves as the designated Level 1 trauma centre for the greater metropolitan area of
Stockholm. This designation requires around-the-clock availability of key services, including neurosurgery, general
surgery, orthopaedic surgery, anaesthesia and intensive care, as well as diagnostic and interventional radiology and
consultants from other relevant specialities. (Källa   från American College of surgeons, plus källa  som säger NKS
har alla dessa saker). Also, the ICU units in Karolinska Universitets sjukhuset are all recognised as Level 3 units
per the Svenska Föreningen för Anestesi och Intensivvård. (källa)

@@

Participants
------------
Included in the study were all adult patients with TBI recorded in the Karolinska trauma registry between 2012 and
2023 who also had been screened for opportunity for improvements. Furthermore, TBI was defined as (a GCS score of ≤8
combined with an Abbreviated Injury Scale ≥3 for the head region ….  ) Patients were eligible irrespective of level <!-- Definition found in Elins paper, same applies?-->
of care and were not restricted to intensive care admissions.

@@

Variables
---------
Outcome

@@

Patient factors and processes

@@

Furthermore, the categorical variables include…
<!-- Choose variables, not sure which ones to chose-->

@@

Logistic regression was used to examine the association between patient and process-related factors and OFI.
Bivariable logistic regression to assess unadjusted associations between patient and process factors and OFI was
initially performed. Variables of clinical relevance and those significant in bivariable analyses were then entered
into a multivariable logistic regression model to estimate adjusted associations. Results are reported as Odds Ratio
(OR) with 95% confidence interval (CI), and a p-value of <0.05 was used as the threshold for statistical
significance. To ensure objectivity, all statistical analyses were initially performed on synthetic data before being
applied to the combined trauma registry and trauma care quality database. Statistical analyses were conducted in R
version x. <!-- I assume we will do both bivariable and multivariable regression? -->

>>>>

Among trauma-related conditions, traumatic brain injury (TBI) is a major cause of death and disability worldwide and
constitutes a significant public health burden.(3) TBI is the result of an external force causing neurological
impairments that may be either transient or permanent.(4) Furthermore, TBI should not be regarded as a single event
but rather as a dynamic process. The initial trauma **often triggers a secondary pathophysiological cascade**, including
ischemia, oedema, and metabolic disturbances, that develop over time and critically influence long-term outcomes. (5)
As such, patients **with TBI** often require admission to the intensive care unit (ICU).

@@

<!-- MGW: Consider discussing the management of trauma first and then the management of TBI, as many of the things you write below also apply to non-TBI trauma patients -->

@@

<!-- MGW: Try to find a balance when it comes to the level of detail that you use. The specific ICD and CPP targets feel a bit too specific. -->

(At Karolinska University Hospital, the setting for this study, similar principles are seen in the Traumahandboken.
For severe TBI, defined as GCS < 8, early airway protection with intubation in unconscious patients, rapid CT
imaging, timely administration of tranexamic acid, and early neurosurgical consultation are recommended. In cases
where there are signs of impending herniation, the handbook advises measures such as controlled hyperventilation and
osmotherapy with mannitol or hypertonic saline. All this is always in conjunction with urgent neurosurgical
evaluation. (Källa Karolinska handboken))  <!-- Remove paragraph? MGW: Yes, I think so.-->

@@

Despite the presence of systems and guidelines in place for these patients, trauma care remains vulnerable to errors,
deviations from established practice, and treatment delays, all of which may negatively influence outcomes. In
response, the American College of Surgeons developed the Trauma Quality Improvement Program (TQIP), a risk-adjusted
benchmarking system designed to assess the quality of trauma care <!-- MGW: I would mention ACS TQIP as one example of a TQIP and refer to the WHO guidelines as more general. I suggest that you start by introducing the concept of TQIP by referring to the WHO guidelines and then mention ACS TQIP as one example -->. Through audit filters and peer review, it
identifies opportunities for improvement (OFI) and ultimately reduces trauma-related mortality (10,11). The WHO, however, insist that the trauma quality improvement programmes must be a continuous, multidisciplinary process. This
entails audit filters, structured morbidity and mortality and preventable death case reviews in order to find
opportunities for improvement (OFI). They also discuss that the formulation of corrective strategies or action plans
in response to identified OFI:s is essential to trauma quality improvement programmes. Furthermore, the WHO
emphasises that TQIP must ensure that corrective measures addressing identified OFI result in tangible and verifiable
improvements in outcomes, thereby confirming that deficiencies are not only recognised but also effectively resolved. (Källa WHO) <!-- add WHO source -->

@@

<!-- MGW: I think this paragraph is good, but I suggest adding a subheading like "Rationale for this study" or something like that. -->

@@

Aim
=====
This study aims to identify and describe opportunities for improvement in the care of adult trauma patients
with TBI, and to examine patient- and process-related factors associated with their occurrence.

@@

Study Design
-----------
A single-centre, registry-based retrospective cohort study was conducted at Karolinska University Hospital in Solna
(NKS). The trauma registry for NKS is a comprehensive trauma registry that **is part of** the Swedish National Trauma
Registry, SweTrau, and is compliant with the Utstein template for uniform reporting after trauma (Källa  Utstein
Template). As such, the registry prospectively records patient demographics, prehospital and emergency department
physiology, times of key interventions such as imaging, surgery and transfer, injury severity scores, and in-hospital outcome.

SweTrau includes patients of all ages who have sustained serious injuries from traumatic events
such as falls, traffic incidents, or other external violence. Since 2024<!-- MGW: Has this changed over the years? Otherwise I suggest that you remove the year. -->, patients are included if a trauma
team activation has occurred, regardless of injury severity, or if they have a New Injury Severity Score (NISS)
greater than fifteen without trauma team activation. Patients transferred to a hospital within seven days of the
traumatic event with a NISS greater than fifteen are also included. Exclusion criteria are cases where the only
traumatic condition is chronic subdural haematoma, trauma alarms without an underlying traumatic event, patients with
protected identity, and patients declared dead on arrival. Since these inclusion and exclusion criteria have changed
over time, patients in this study will be filtered to ensure they fulfil the criteria relevant to the study period
(Källa SweTrau årsbok 2024). <!--Use the 2023 version instead? Data used is up to 2023 -->

@@

Setting
---------
Karolinska University Hospital in Solna houses Trauma Centre Karolinska (TCK), Sweden’s largest provider of advanced
trauma care for both adults and children<!-- MGW: Do you have a source for this? -->. The trauma center has around-the-clock availability of key services, including neurosurgery, general
surgery, orthopaedic surgery, anaesthesia and intensive care, as well as diagnostic and interventional radiology and
consultants from other relevant specialities. (Källa   från American College of surgeons, plus källa  som säger NKS
har alla dessa saker). Also, the ICU units in Karolinska Universitets sjukhuset are all recognised as Level 3 units
according to the **Swedish Society of Anaesthesiology and Intensive Care Medicine**. (källa)

<!-- The numbers below belong to the results section, so they should be moved there. -->

@@

Participants
------------
All adult patients with TBI recorded in the Karolinska trauma registry between 2012 and
2023 who also had been screened for opportunity for improvements **were included in this study**. Furthermore, TBI was defined as (a GCS score of ≤8
combined with an Abbreviated Injury Scale ≥3 for the head region ….  ) Patients were eligible irrespective of level <!-- Definition found in Elins paper, same applies?--><!-- Yes -->
of care.

@@

Variables
---------

### Outcome

@@

### Patient factors and processes

@@

Furthermore, the categorical variables include…
<!-- Choose variables, not sure which ones to chose--> 
<!-- MGW: Start with the ones in Husseins and Elins papers-->

@@

<!-- MGW: Regardless of the analysis method, we need to handle missing data somehow. You can use complete case analysis, which means that you only include cases with complete data for all variables of interest. -->

Logistic regression was used to examine the association between patient and process-related factors and OFI.
Bivariable logistic regression to assess unadjusted associations between patient and process factors and OFI was
initially performed. Variables of clinical relevance and those significant in bivariable analyses were then entered
into a multivariable logistic regression model to estimate adjusted associations. <!-- MGW: I think you should decide which ones to include in the multivariable model from the start, regardless of their significance in the bivariable analysis --> Results are reported as Odds Ratio
(OR) with 95% confidence interval (CI), and a p-value of <0.05 was used as the threshold for statistical
significance. To ensure objectivity, all statistical analyses were initially performed on synthetic data before being
applied to the combined trauma registry and trauma care quality database. Statistical analyses were conducted in R
version x. <!-- I assume we will do both bivariable and multivariable regression? -->
